A Handbook for Tenured Track Faculty in Computer Science in the US

### 2024-09-30 faculty.tex
updates

\section{Most people suck at grant writing (at first)} Most new assistant professors have no experience in grant writing, and struggle with it in the beginning.  It becomes especially stressful starting your 3rd year when departmental evaluation starts to give you more pressure and warning if you don't have enough funding. By your midterm or fourth-year review, you are expected to have at least some external funding, and can be in serious trouble if you don't have any. So start early, and don't wait until the last minute to start writing grants.  It's a skill that can be learned, and you will get better over time.  The key is to keep trying, and to not give up.  If you are not getting rejected, you are not aiming high enough.  The best researchers in the world get rejected all the time.  It's part of the process.  So don't get discouraged.  Keep trying, and you will get better over time.

@@

Most depts. would allow faculty to pick their favorite courses to teach.  In some cases, they might ask you to teach an existing core undergraduate courses (e.g., Theory, Operating Systems) and develop your own grad course (e.g., Software Verification).
It is likely that you get to repeat the same courses. This significant reduces the amount of time you need to prepare for the course.  However, you should not teach the same course every semester, and you will want to alternate between grad and undergrad courses.

\begin{commentbox}[Vu]
    At GMU, I teach the same Software Specification and Construction course at both grad (SWE-619) and undergrad (SWE-419) levels. This is very nice because I don't have to prepare much for the course.  I did try to teach an AI Assurance graduate seminar once, but never liked it much because it only has MS students, who were not very interested in research.

    In contrast, I teach various courses at UNL. I teach the several undergraduate courses including Theory, Software Engineering, and design my own two courses on Compilers and Software Analysis. I also create a grad seminar on Software Verification.

    I feel that UNL is a bit more flexible in creating new classes compared to GMU, where I am pretty much the main instructor for SWE-619/SWE-419.
  \end{commentbox}

\section{Teaching Evaluation}

At the end of each semester, students are asked to evaluate the instructor. The results are shared anonymously with the instructor to help them improve their teaching. Teaching evaluation is also commonly used in annual and tenure evaluation.

@@

\begin{commentbox}[Vu]
A strategy that I found effective is that the whole class gets a small bonus point if everyone completes the evaluation. The bonus is calculated as the percentage of responses multiplied by a factor (X). So, if X is 0.5 and everyone participates, they get 0.5 points each. If only half respond, it's 0.25 points each. It might seem small, but it can make a difference for borderline grades. Since I don't curve grades, this is their only chance to boost their scores. Many students have benefited from this in my previous classes, which really grabs their attention. Plus, since evaluations happen before finals and grades aren't finalized, everyone's motivated to participate—they don't want to miss out. Students not only do it themselves but also remind their friends. It's been working well for me: Spring '24 I had a 90\% response rate (25 out of 28 students), and Fall '23 was a perfect 100\% (16 out of 16 students). Before, my response rate was about 50\%, which is already higher than the average.
\end{commentbox}

@@

@@

Successfully mentoring students is a key and most rewarding part of your job as a faculty member. 
There will be times when you feel discouraged and frustrated, but there will also be times when you feel proud and happy.  
For P\&T you will need to successfully mentoring students.  Many universities expect you to need to graduate at least one PhD student to get tenure.

@@

I have worked with many undergrads (15+ since 2016). My success rate (publish a research paper with me) with undergrads

\section{Mentoring Postdocs}
\section{Mentoring Junior Faculty}

@@

After 5 or 6 years, it is time for you to go up (or apply) for tenure.  You will need to submit a tenure package.  The department will then review your package and vote on whether to recommend you for tenure.  If the department votes in your favor, the package will then go to the college, and then to the university (e.g., provost, president, board of visitors, etc).  The university will then make the final decision on whether to grant you tenure (typically they follow the department's recommendation, but not always).

@@

\section{Change in Department. Moving from UNL to GMU}

I joined UNL in 2016 and was very happy there until 2020, when people start leaving and the dept. changed its leadership and organization.  All of these uncertainties and other issues lead to my decision to be on the job market again.

I will skip the details and just summarize several important and disappointing events.

\begin{itemize}
\item 2019: four senior professors (in SE) left.  One of these 4 was the dept chair (and he left just after 2 months of being chair). There were issues of upper admin didn't work well with the dept chair (and this issue has been going for years).
\item 2020: College of Engr pressure dept to move entirely to it. The new chair candidate only agrees to accept the job if the faculty moved to CoE, putting pressure on the move. Faculty are split on the decision and eventually moved with lots of mistrusts and concerns under CoE.
\item Salary inversion:  new chair offers new AP salary that is much higher than current AP, leading to jokes that current AP should just reapply to UNL.  This became a serious issue and many people at UNL use this as a reason on why they move. 
\item Senior faculty fighting on mailing list.

\end{itemize}

Note that even after leaving Lincoln, my family and I still have a soft spot for the city and its people.  It was a great place to live and raise a family, but unfortunately not a good place to work for me.

>>>>

\section{Most people suck at grant writing (at first)} Most new assistant professors have no experience in grant writing, and struggle with it. It becomes especially stressful starting your 3rd year when departmental evaluation starts to look at your funding. By your midterm or fourth-year review, you are expected to have at least some external funding, and can be in trouble if you don't have any. However, despite initial struggles, most people get better at grant writing over time.  The best researchers in the world get rejected all the time. So don't get discouraged.  Keep trying, and you will get better over time.

@@

Most depts. would allow faculty to pick their favorite courses to teach. In some cases, they let  you develop your own grad course (e.g., Software Verification) but ask you to teach an existing core undergraduate courses (e.g., Theory, Operating Systems).
It is likely that you get to repeat the same courses. This significant reduces the amount of time you need to prepare for the course.  However, you should not teach the same course every semester, and you will want to alternate between grad and undergrad courses.

\begin{commentbox}[Vu]
    At GMU, I teach the same Software Specification and Construction course at both grad (SWE-619) and undergrad (SWE-419) levels. This is very nice because I don't have to prepare much for the course.  I did try to teach an AI Assurance graduate seminar once, but never liked it much because it only has MS students, who were not very interested in research.
    \\

    In contrast, I teach various courses at UNL. I teach the several undergraduate courses including Theory, Software Engineering, and design my own two courses on Compilers and Software Analysis. I also create a grad seminar on Software Verification.\\

    I feel that UNL is a bit more flexible in creating new classes compared to GMU.  However, I spent much less time prepare for courses at GMU.
  \end{commentbox}

\paragraph{Co-Teaching}
Some places encourage new TT faculty to co-teach with another faculty. This is a way to share the workload and to learn from your more experienced colleague. It has many benefits, but it can also be challenging if you and your co-teacher have different teaching styles.

\begin{commentbox}[Vu]
I co-taught an undergrad Software Engineering course once at UNL with a teaching faculty.  I didn't like it because the teaching faculty was very experienced and good at teaching, and I felt that I was not contributing much.  I never co-taught again after that.
\end{commentbox}

\section{Teaching Evaluation}\label{sec:teaching-evaluation}

Teaching evaluations are a critical component of your annual review and P\&T. It provides an analysis on how students and peers perceive your teaching. 
You pay attention to evaluations, and make changes to your teaching based on the feedback, and show that you continuously strive to improve your teaching.

\paragraph{Student Evaluation}
At the end of each semester, students are asked to evaluate the instructor. The evaluation often includes both quantitative (e.g., ratings on a scale of 1-5) and qualitative (e.g., comments) feedback.  In addition to default questions, you can add your own questions. The results are shared anonymously with the instructor to help them improve their teaching.

@@

\begin{commentbox}[Vu]
  A strategy that I found effective is that the whole class gets a small bonus point if everyone completes the evaluation. The bonus is calculated as the percentage of responses multiplied by a factor (X). So, if X is 0.5 and everyone participates, they get 0.5 points each. If only half respond, it's 0.25 points each. It might seem small, but it can make a difference for borderline grades. Since I don't curve grades, this is their only chance to boost their scores. Many students have benefited from this in my previous classes, which really grabs their attention. Plus, since evaluations happen before finals and grades aren't finalized, everyone's motivated to participate—they don't want to miss out. Students not only do it themselves but also remind their friends. 
  \\

  It's been working well for me: Spring '24 I had a 90\% response rate (25 out of 28 students), and Fall '23 was a perfect 100\% (16 out of 16 students). Before, my response rate was about 50\%, which is already higher than the average.
  \end{commentbox}

\paragraph{Peer Evaluations}

In addition to student feedback, peer evaluations can offer valuable insights on your teaching. At many places, peer evaluations are required for P\&T.
Request evaluations from experienced colleagues who can observe your class and provide constructive criticism. Like with student evaluation, incorporating these peer feedback into your teaching and describe your changes and improvements.  Peer evaluation letters can also help you get Teaching Awards and other recognitions.

@@

\begin{commentbox}[Vu]
  I have recruited my first PhD student Guolong from my seminar and many undergrads from my classes. However, other than Guolong, the undergrads recruited from my class did not work out well.  I found that undergrads recruited from my class are not as motivated as those who come to me on their own. So, I stop recruiting students from my class and only take those who come to me on their own. More about recruiting students in \autoref{sec:recruiting}
\end{commentbox}

@@

Successfully mentoring students is a key and most rewarding part of your job as a faculty member. 
There will be times when you feel discouraged and frustrated, but there will also be times when you feel proud and happy.

For P\&T you will need to successfully mentoring students.  Many universities expect you to need to graduate at least one PhD student to get tenure.

\section{Recruiting}\label{sec:recruiting}

@@

I have worked with many undergrads (15+ since 2016).

\section{Mentoring Postdocs}

I never formally mentored a postdoc, but I have worked with several postdocs. They are more like colleague than students as they are more mature and independent, e.g., writing own papers and leading projects.

\section{Mentoring Junior Faculty}

@@

After 5 or 6 years, it is time for you to go up (apply) for tenure.  You will need to submit a tenure package.  The department will then review your package and vote on whether to recommend you for tenure.  If the department votes in your favor, the package will then go to the college, and then to the university (e.g., provost, president, board of visitors, etc).  The university will then make the final decision on whether to grant you tenure (typically they follow the department's recommendation, but not always).

@@

% \section{Change in Department. Moving from UNL to GMU}

% I joined UNL in 2016 and was very happy there until 2020, when people start leaving and the dept. changed its leadership and organization.  All of these uncertainties and other issues lead to my decision to be on the job market again.

% I will skip the details and just summarize several important and disappointing events.

% \begin{itemize}
% \item 2019: four senior professors (in SE) left.  One of these 4 was the dept chair (and he left just after 2 months of being chair). There were issues of upper admin didn't work well with the dept chair (and this issue has been going for years).
% \item 2020: College of Engr pressure dept to move entirely to it. The new chair candidate only agrees to accept the job if the faculty moved to CoE, putting pressure on the move. Faculty are split on the decision and eventually moved with lots of mistrusts and concerns under CoE.
% \item Salary inversion:  new chair offers new AP salary that is much higher than current AP, leading to jokes that current AP should just reapply to UNL.  This became a serious issue and many people at UNL use this as a reason on why they move. 
% \item Senior faculty fighting on mailing list.

% \end{itemize}

% Note that even after leaving Lincoln, my family and I still have a soft spot for the city and its people.  It was a great place to live and raise a family, but unfortunately not a good place to work for me.
